---
title: "String Practice"
author: "Lewis White"
date: "2022-08-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


With your cursor in your code chunk, go up to Addins in the top bar of RStudio. From the drop-down menu, choose ‘Paste as data frame’. Make sure to add code to store the data frame as alarm_report


```{r}
alarm_report <- tibble::tribble(
         ~date,           ~building,                                              ~alarm_message,
  "2020-03-14", "Engineering-North", "10:02am – HVAC system down, facilities management alerted",
  "2020-03-15",         "Bren Hall",                    "8:24am – Elevator North out of service",
  "2020-04-10", "Engineering-South",     "12:41am – Fire alarm, UCSB fire responded and cleared",
  "2020-04-18",        "Engr-North",     "9:58pm – Campus point emergency siren, UCPD responded"
  )

```

Replace the “Engr” with “Engineering” in the building column

Separate the building column into two separate columns, building and wing, separated at the dash

Separate the message time from the rest of the message by separating at --

```{r}
alarm_report_new <- alarm_report %>%
  mutate(building = str_replace(building, pattern = "Engineering", replacement = "Engr")) %>% 
  separate(col = building, into = c("building", "wing"), sep = "-") %>%
  filter(str_detect(alarm_message, "responded")) %>%
  separate(col = alarm_message, into = c("alarm_time", "alarm_message"), sep = "–")
  
```

Convert the date column to a Date class using lubridate

```{r}
alarm_report_new_date <- alarm_report_new %>%
  mutate(date = lubridate::ymd(date))
```

